---
title: "CCS_MainFig2"
output: html_document
date: "2023-11-27"
editor_options: 
  chunk_output_type: console
---
## Install packages
```{r}
library(devtools)
install_github("jokergoo/ComplexHeatmap")
```

# Load libraries
```{r}
library(ComplexHeatmap)
```

# Find all markers - distinct from CC
```{r}
Idents(ccs) <- ccs$all
markers_ccs_onehis <- FindAllMarkers(ccs, logfc.threshold = 0)

markers_ccs_cc <- # To find CCS specific markers
  FindMarkers(
    ccs,
    ident.1 = c(
      "SA node",
      "AV node",
      "Lower nodal bundle",
      "His bundle",
      "Proximal bundle branch",
      "Distal bundle branch",
      "Purkinje fibers",
      "Proliferative CCS"
    ),
    ident.2 = "Contractile cardiomyocytes",
    logfc.threshold = 0
  )
```

```{r}
ccs_ccsonly <- subset(
  ccs,
  idents = c(
    "SA node",
    "Compact AV node",
    "Lower nodal bundle",
    "His bundle",
    "Proximal bundle branch",
    "Distal bundle branch",
    "Purkinje fibers",
    "Proliferative CCS"
  )
)

ccs_ccsonly <- NormalizeData(ccs_ccsonly)
ccs_ccsonly <- FindVariableFeatures(ccs_ccsonly, nfeatures = 5000)
ccs_ccsonly <-
  ScaleData(
    ccs_ccsonly,
    vars.to.regress = c('nCount_RNA', 'percent.mito'),
    features = rownames(ccs_ccsonly)
  )
ccs_ccsonly <- RunPCA(ccs_ccsonly, npcs = 50)
fig <- ElbowPlot(ccs_ccsonly, ndims = 50)


markers_ccs_ccsonly <-
  FindAllMarkers(ccs_ccsonly, logfc.threshold = 0)
write.table(
  markers_ccs_ccsonly,
  file = "~/markers_within_ccs_only.txt",
  sep = '\t',
  col.names = NA
)

rm(markers_ccs_ccsonly)

```

# list of significant genes
```{r}
markers_ccsonly_basic_onehis <- markers_ccs_ccsonly %>%
  filter(p_val_adj <= 0.05 & avg_log2FC >= 1) %>% # for FC > 2
  group_by(cluster) %>%
  top_n(50, avg_log2FC) %>%
  pull(gene)
```

# Heatmap - Used ComplexHeatmap to show duplicates
```{r}
scaled_markers_ccsonly_basic_onehis <-
  ccs_ccsonly[["RNA"]]@scale.data[markers_ccsonly_basic_onehis,] %>% as.matrix()
cluster_anno <- ccs_ccsonly@meta.data$sub_all_onehis
```

# Decide on color palette
```{r}
#show_col(colorRampPalette(brewer.pal(9,"Reds"))(100))
#myCol <- colorRampPalette(brewer.pal(9,"Reds"))(50)
myBreaks <- seq(-2, 2.8, length.out = 50)
myCol <- colorRampPalette(c("dodgerblue", "black", "yellow"))(50)
```

```{r}
tiff(
  "~/fig2a_sc_heatmap_ccsonly_log2fc1_pval0.05p.tiff",
  width = 10,
  height = 6.2,
  units = "in",
  res = 500
)

fig <- Heatmap(
  scaled_markers_ccsonly_basic_onehis,
  name = "Scaled\nExpression",
  column_split = factor(cluster_anno),
  cluster_columns = FALSE,
  show_column_dend = FALSE,
  cluster_column_slices = TRUE,
  column_gap = unit(0.5, "mm"),
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = circlize::colorRamp2(myBreaks, myCol),
  row_names_gp = gpar(fontsize = 4),
  top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = mycolors2))),
  #column_title_rot = 90,
  column_title_gp = gpar(fontsize = 0),
  show_column_names = FALSE,
  use_raster = TRUE
)

draw(fig)

dev.off()
```

# Fig2a (Left, Below) - Make a zoomed-in heatmap for proximal part of CCS only 
```{r}
subset_proxccsonlyheatmap <-
  subset(
    ccs_ccsonly,
    idents = c(
      "SA node",
      "Compact AV node",
      "Lower nodal bundle",
      "His bundle",
      "Proximal bundle branch"
    )
  )
```

# Do the same on smaller population
```{r}
scaled_markers_prox_ccsonly_forheatmap_onehis <-
  subset_proxccsonlyheatmap[["RNA"]]@scale.data[markers_ccsonly_basic_onehis, ] %>% as.matrix()
cluster_anno <- subset_proxccsonlyheatmap@meta.data$sub_all_onehis
```

```{r}
tiff("~/fig2a_sc_heatmap_ccsonly_prox_log2fc1_pval0.05.tiff", width = 10, height = 2, units = "in", res = 500)

fig <- Heatmap(scaled_markers_prox_ccsonly_forheatmap_onehis, 
        name = "Scaled\nExpression",  
        column_split = factor(cluster_anno),
        cluster_columns = FALSE,
        show_column_dend = FALSE,
        cluster_column_slices = TRUE,
        column_gap = unit(0.5, "mm"),
        cluster_rows = FALSE,
        show_row_dend = FALSE,
        col = circlize::colorRamp2(myBreaks, myCol),
        row_names_gp = gpar(fontsize = 4),
        top_annotation = HeatmapAnnotation(foo = anno_block(gp = gpar(fill = mycolors2))),
        #column_title_rot = 90,
        column_title_gp = gpar(fontsize = 0),
        show_column_names = FALSE,
        use_raster = TRUE)

draw(fig)

dev.off()
```


# Fig2A - Spatial heatmap - make matrix
```{r}
spatialheatmap <-
  read.table(
    "~/spatial_fc1.5_ccs_correspondingcntrl.txt"
  )
spatialheatmap$V1 <- make.unique(spatialheatmap$V1)
row.names(spatialheatmap) <- spatialheatmap$V1
spatialheatmap$V1 <- NULL

spatialheatmap_labels <-
  read.table(
    "~/q3_ExpressionFile_wlnb_forfig2_heatmap_groups_ccsonly.txt"
  )

colnames(spatialheatmap) <- spatialheatmap_labels
```

# Scale data
```{r}
scaled_spatialheatmap <- as.matrix(t(scale(t(spatialheatmap))))
```

# Color palette - blue
```{r}
colorRampPalette(brewer.pal(9,"Blues"))(100)
quantile(scaled_spatialheatmap, c(0.1, 0.95))
heatmap_spatial_col = circlize::colorRamp2(c(-1, 0, 2), c("#F7FBFF", "#6DAFD6", "#08306B")) # used for FC>1.2, pval<0.05
```

# Fig2A - Spatial heatmap
```{r}
tiff("~//fig2a_spatial_fc1.5_vs_correspondingcontrols.tiff", width = 4, height = 5, units = "in", res = 500)


fig <- Heatmap(
  scaled_spatialheatmap,
  name = "Scaled\nExpression",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = heatmap_spatial_col,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 4)
)

draw(fig)

dev.off()
```




# Fig2B - Find marker genes for grouped clusters
```{r}
markers_nodes <-
  FindMarkers(
    ccs_ccsonly,
    ident.1 = c("SA node", "Compact AV node"),
    ident.2 = c(
      "Lower nodal bundle",
      "His bundle",
      "Proximal bundle branch",
      "Distal bundle branch",
      "Purkinje fibers",
      "Proliferative CCS"
    ),
    logfc.threshold = 0
  )

write.table(markers_nodes,
            file = "~/fig2b_markers_sc_nodes_within_ccs_only.txt",
            sep = '\t',
            col.names = NA)

markers_proxvcs <-
  FindMarkers(
    ccs_ccsonly,
    ident.1 = c("Lower nodal bundle",
                "His bundle",
                "Proximal bundle branch"),
    ident.2 = c(
      "SA node",
      "Compact AV node",
      "Distal bundle branch",
      "Purkinje fibers",
      "Proliferative CCS"
    ),
    logfc.threshold = 0
  )

write.table(markers_proxvcs,
            file = "~/fig2b_markers_sc_proxvcs_within_ccs_only.txt",
            sep = '\t',
            col.names = NA)

markers_distvcs <-
  FindMarkers(
    ccs_ccsonly,
    ident.1 = c("Distal bundle branch",
                "Purkinje fibers"),
    ident.2 = c(
      "SA node",
      "Compact AV node",
      "Lower nodal bundle",
      "His bundle",
      "Proximal bundle branch",
      "Proliferative CCS"
    ),
    logfc.threshold = 0
  )

write.table(markers_distvcs,
            file = "~/fig2b_markers_sc_distvcs_within_ccs_only.txt",
            sep = '\t',
            col.names = NA)

```

# Fig2B - Dotplot for scRNAseq
```{r}
dotplot_genes <-
  read.table(
    "~/fig2b_spatial_grouped_genelist.txt"
  )
dotplot_genes <- dotplot_genes$V1
```

```{r}
DotPlot(
  ccs,
  features = dotplot_genes,
  #idents = rev(levels(ccs)),
  cols = c('whitesmoke', 'deeppink4'),
  col.min = 0,
  col.max = 2,
  dot.scale = 10,
  dot.min = -0.10
)  &
  guides(color = guide_colorbar(title = 'Average\nExpression')) &
  guides(size = guide_legend(title = 'Percent\nExpressed')) &
  RotatedAxis() &
  scale_size(breaks = c(0, 20, 40, 60, 80, 100)) &
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
      face = 'italic',
      size = 12,
      angle = 90,
      vjust = 0.5,
      hjust = 0.5
    ),
    axis.ticks.length.x.bottom = unit(3, 'mm'),
    axis.ticks.length.y.left = unit(3, 'mm'),
    axis.text.y = ggtext::element_markdown(size = 13,
                                           colour = rev(mycolors2)),
    legend.position = "right",
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.key.width = unit(0.2, 'cm'),
    legend.key.height = unit(0.2, 'cm')
  )
```

# Fig2B - Heatmap for spatial (grouped clusters)
```{r}
fig2b_spatialheatmap <-
  read.table(
    "~fig2b_spatial_grouped_wcntrls.txt"
  )

row.names(fig2b_spatialheatmap) <- fig2b_spatialheatmap$V1
fig2b_spatialheatmap$V1 <- NULL
spatialheatmap_labels <-
  read.table(
    "~/q3_ExpressionFile_wlnb_forfig2_heatmap_groups.txt"
  )
colnames(fig2b_spatialheatmap) <- spatialheatmap_labels
```

# Scale data
```{r}
scaled_fig2_spatialheatmap <-
  as.matrix(t(scale(t(
    fig2b_spatialheatmap
  ))))
scaled_fig2_spatialheatmap <- t(scaled_fig2_spatialheatmap)
```

# Color palette - blue
```{r}
colorRampPalette(brewer.pal(9,"Blues"))(100)
heatmap_spatial_col = circlize::colorRamp2(c(-1.5, 0, 2), c("#F7FBFF", "#6DAFD6", "#08306B")) # used for FC>1.2, pval<0.05
```

# Fig2A - Spatial heatmap
```{r}
tiff("~/fig2b_groupedclusters_spatialheatmap_wcntrls.tiff", width = 10, height = 3, units = "in", res = 500)

fig <- Heatmap(
  scaled_fig2_spatialheatmap,
  name = "Scaled\nExpression",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = heatmap_spatial_col,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6)
)

draw(fig)

dev.off()
```
