---
title: "CCS_MainFig4"
output: html_document
date: "2023-09-15"
editor_options: 
  chunk_output_type: console
---
# Color palette
```{r}
mycolors_goodyer <- c(
  "#661100", #E16.5 - SAN-Head
  "#990000", #E16.5 - SAN-Tail
  "#B80F0A", #E16.5 - Tz-SAN
  "#DD561C", #E16.5 - cAVN
  "#f7a92b", #E16.5 - Nodal AVR
  "#F98228", #E16.5 - Atrial TZ
  "#FADA5E", #E16.5 - His bundle
  "#5AAE61", #E16.5 - Ventricular TZ
  "#2166AC", #E16.5 - Tz-PF
  "#512889" #E16.5 - PF
)


mycolors2 <- c(    
  "#4d0000", #P - SA node
  "#E8601C", #P - AV node
  "#F1932D", #P - Lower nodal bundle
  "#F6C141", #P - His bundle
  "#3DBF3D", #P - Proximal bundle branches
  "#77C9E9", #P - Distal bundle branches
  "#7c7cc1", #P - Purkinje fibers
  "#4C4C4C", #P - Proliferative CCS
  "#BBBBBB" #P - Contractile cardiomyocytes
)


mycolors_ohgoodyer <-
  c(
    "#661100", #E16.5 - SAN-Head
    "#990000", #E16.5 - SAN-Tail
    "#B80F0A", #E16.5 - Tz-SAN
    "#DD561C", #E16.5 - cAVN
    "#F98228", #E16.5 - Atrial TZ
    "#f7a92b", #E16.5 - Nodal AVR
    "#FADA5E", #E16.5 - His bundle
    "#5AAE61", #E16.5 - Ventricular TZ
    "#2166AC", #E16.5 - Tz-PF
    "#512889", #E16.5 - PF
    "#4d0000", #P - SA node
    "#E8601C", #P - Compact AV node
    "#F1932D", #P - Lower nodal bundle
    "#F6C141", #P - His bundle
    "#3DBF3D", #P - Proximal bundle branches
    "#77C9E9", #P - Distal bundle branches
    "#7c7cc1", #P - Purkinje fibers
    "#4C4C4C", #P - Proliferative CCS
    "#BBBBBB" #P - Contractile Myocytes
  )
```

# Fig 4A - UMAP for Embryonic CCS
```{r}
DimPlot(
  goodyer_ccs,
  group.by = 'goodyer',
  cols = mycolors_goodyer,
  pt.size = 1.5
) &
  NoAxes() & NoLegend() &
  theme(text = element_text(size = 15),
        plot.title = element_blank(),
        panel.background = element_rect(fill = "transparent", colour = NA),  
        plot.background = element_rect(fill = "transparent", colour = NA))

```

# Fig 4A - UMAP for Postnatal CCS
```{r}
DimPlot(
  ccs,
  cols = mycolors2,
  pt.size = 0.5
) &
  NoAxes() &

  theme(text = element_text(size = 15),
        plot.title = element_blank(),
        panel.background = element_rect(fill = "transparent", colour = NA),  
        plot.background = element_rect(fill = "transparent", colour = NA)) &
  NoLegend()
```

# Fig 4A - Integrated CCS - Postnatal only
```{r}
nacol <- "grey92"
Idents(ohgoodyer_int) <- ohgoodyer_int$postnatal

DimPlot(
  ohgoodyer_int,
  cols = mycolors2,
  pt.size = 0.8,
  order = T,
  shuffle = T,
  na.value = alpha(nacol, 0.1),
) &
  NoAxes() &
  theme(text = element_text(size = 15),
        plot.title = element_blank(),
        panel.background = element_rect(fill = "transparent", colour = NA),  
        plot.background = element_rect(fill = "transparent", colour = NA))
```

# Fig 4A - Integrated CCS - Embryonic only
```{r}
nacol <- "grey92"
Idents(ohgoodyer_int) <- ohgoodyer_int$embryonic

DimPlot(
  ohgoodyer_int,
    cols = mycolors_goodyer,
  na.value = alpha(nacol, 0.1),
  order = T,
  shuffle = T,
  pt.size = 0.8
) &
  NoAxes() & 
  theme(
    text = element_text(size = 15),
    plot.title = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
  )

```

# Figure 4B - Correlation plot - ohgoodyer_int
```{r}
library(corrplot)

DefaultAssay(ohgoodyer_int) <- "integrated"

avg_ohgoodyer_int <-
  AverageExpression(ohgoodyer_int, return.seurat = T) # make averaged seurat object
avg_ohgoodyer_int_mtx <-
  avg_ohgoodyer_int@assays[["integrated"]]@data # retrieve normalized exp
avg_ohgoodyer_int_mtx_cor <-
  cor(avg_ohgoodyer_int_mtx) # run correlation

write.table(
  avg_ohgoodyer_int_mtx_cor,
  "~fig4b_1_correlation.txt",
  sep = '\t',
  col.names = NA
)
```

# After re-organizing correlation values, re-read & scale data
```{r}
avg_ohgoodyer_int_mtx_cor <-
  read.table("~fig4b_2_correlation_reorg.txt",
             header = T)
rownames(avg_ohgoodyer_int_mtx_cor) <- avg_ohgoodyer_int_mtx_cor$X
avg_ohgoodyer_int_mtx_cor$X <- NULL

avg_ohgoodyer_int_mtx_cor <- scale(t(avg_ohgoodyer_int_mtx_cor))
avg_ohgoodyer_int_mtx_cor <- t(avg_ohgoodyer_int_mtx_cor)
```

```{r}
label_sc <-
  read.table(
    "/Users/yenaoh/Desktop/Mission_PhD/Paper/CCS_Manuscript/1_NatCommSubmission/1_R2R/1_R2R_fig/fig4/onehis/label.txt"
  )
label_sc <- label_sc$V1
rownames(avg_ohgoodyer_int_mtx_cor) <- label_sc

write.table(
  avg_ohgoodyer_int_mtx_cor,
  "/Users/yenaoh/Desktop/Mission_PhD/Paper/CCS_Manuscript/1_NatCommSubmission/1_R2R/1_R2R_fig/fig4/onehis/fig4b_3_correlation_scaled.txt",
  sep = '\t',
  col.names = NA
)
```

```{r}
# colorRampPalette(brewer.pal(3,"YlOrRd"))(100)
heatmap_corr_p_e = circlize::colorRamp2(c(-1.5,0,1.5), colors = c("#ffffc0", "#fd8d3c", "#800026"))
```

```{r}
tiff("~/fig4b_embryonic_postnatal_corr_heatmap_nodendro.tiff", width = 3.5, height = 4.5, units = "in", res = 500)

fig <- Heatmap(
  avg_ohgoodyer_int_mtx_cor,
  name = "Scaled\nCorrelation",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_column_dend = TRUE,
  col = heatmap_corr_p_e,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 6),
  column_names_gp = gpar(fontsize = 6)
)

draw(fig)

dev.off()
```

# Fig 4C - DotPlot of common genes
```{r}
markers_common_notcc <-
  read.table("~/fig4c_commonmarkers_notCC_20240205.txt")
markers_common_notcc <- markers_common_notcc$V1
markers_common_notcc <- unique(markers_common_notcc)
markers_common_notcc <- rev(markers_common_notcc)
```

```{r}
DotPlot(
  goodyer_ccs,
  assay = "RNA",
  features = markers_common_notcc,
  cols = c('whitesmoke', 'deeppink4'),
  #dot.scale = 5,
  col.min = -0.5,
  col.max = 2
)  & coord_flip() &
  guides(color = guide_colorbar(title = 'Average\nExpression')) &
  guides(size = guide_legend(title = 'Percent\nExpressed')) &
  RotatedAxis() &
  scale_size(breaks = c(0, 25, 50, 75, 100)) &
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
      size = 10,
      angle = 45,
      vjust = 1,
      hjust = 1
    ),
    axis.ticks.length.x.bottom = unit(1, 'mm'),
    axis.text.y = ggtext::element_markdown(size = 13),
    legend.position = "right",
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.width = unit(0.35, 'cm'),
    legend.key.height = unit(0.35, 'cm')
  )

```

# Fig 4D - VlnPlot of different genes - Embryonic vs Postnatal
```{r}
DefaultAssay(ohgoodyer_int) <- "RNA"

markers_ohgoodyer_sant <-
  FindMarkers(
    ohgoodyer_int,
    ident.1 = "E16.5 - SAN-Tail",
    ident.2 = "P - SA node",
    logfc.threshold = 0
  )
write.table(
  markers_ohgoodyer_sant,
  "~/fig4d_ohgoodyer_markers_ESANtail_PSAN_RNA.csv",
  sep = ',',
  col.names = NA
)

markers_ohgoodyer_avn <-
  FindMarkers(
    ohgoodyer_int,
    ident.1 = "E16.5 - Compact AVN",
    ident.2 = "P - AV node",
    logfc.threshold = 0
  )
write.table(
  markers_ohgoodyer_avn,
  "~/fig4d_ohgoodyer_markers_EcAVN_PAVN_RNA.csv",
  sep = ',',
  col.names = NA
)

markers_ohgoodyer_his <-
  FindMarkers(
    ohgoodyer_int,
    ident.1 = "E16.5 - His bundle",
    ident.2 = "P - His bundle",
    logfc.threshold = 0
  )
write.table(
  markers_ohgoodyer_his,
  "~/fig4d_ohgoodyer_markers_EHis_PHis_RNA.csv",
  sep = ',',
  col.names = NA
)

markers_ohgoodyer_pf <-
  FindMarkers(
    ohgoodyer_int,
    ident.1 = "E16.5 - PF",
    ident.2 = "P - Purkinje fibers",
    logfc.threshold = 0
  )
write.table(
  markers_ohgoodyer_pf,
  "~/fig4d_ohgoodyer_markers_EPF_PPF_RNA.csv",
  sep = ',',
  col.names = NA
)
```

# Find top markers
```{r}
markers_ohgoodyer_all_top_10pct <- markers_ohgoodyer_all %>%
  filter(p_val_adj <= 0.05 & avg_log2FC >= 1) %>%
  filter(max(pct.1, pct.2) >= 0.1) %>%
  group_by(cluster) %>%
  top_n(10, avg_log2FC) %>%
  pull(gene)
```

```{r}
markers_diff_san <-
  read.table("~/fig4d_1_san_embvspost_RNA_n15.txt")
markers_diff_san <- markers_diff_san$V1

markers_diff_avn <-
  read.table("~/fig4d_2_avn_embvspost_RNA_n15.txt")
markers_diff_avn <- markers_diff_avn$V1

markers_diff_his <-
  read.table("~/fig4d_3_his_embvspost_RNA_n15.txt")
markers_diff_his <- markers_diff_his$V1

markers_diff_pf <-
  read.table("~/fig4d_4_pf_embvspost_RNA_n15.txt")
markers_diff_pf <- markers_diff_pf$V1
```

# Diff genes - SAN
```{r}
VlnPlot(
     ohgoodyer_int,
     assay = "RNA",
     features = markers_diff_san,
     idents = c("P - SA node", "E16.5 - SAN-Tail"),
     fill.by = 'ident',
     stack = TRUE,
     flip = T,
     cols = c("#990000","#990000")
 ) & stat_summary(fun = mean, geom='crossbar', width = 0.2, colour = "black") &
     theme(
         axis.title.x = element_blank(),
         axis.text.x = element_blank(),
         # axis.text.x = element_text(
         #     size = 16,
         #     angle = 40,
         #     vjust = 1,
         #     hjust = 1
         # ),
         axis.ticks.length.x = unit(0.1, "cm"),
         axis.ticks.length.y = unit(0.1, "cm"),
         axis.title.y = element_blank(),
         axis.text.y = element_text(size = 12),
         strip.text = element_text(size = 15, hjust = 0, face = "italic")
     ) &
     NoLegend()
```

# Diff genes - AVN
```{r}
VlnPlot(
  ohgoodyer_int,
  assay = "RNA",
  features = markers_diff_avn,
  idents = c("P - AV node", "E16.5 - Compact AVN"),
  fill.by = 'ident',
  stack = TRUE,
  flip = T,
  cols = c("#F98228", "#F98228")
) &
  stat_summary(
    fun = mean,
    geom = 'crossbar',
    width = 0.2,
    colour = "black"
  ) &
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    # axis.text.x = element_text(
    #     size = 16,
    #     angle = 40,
    #     vjust = 1,
    #     hjust = 1
    # ),
    axis.ticks.length.x = unit(0.1, "cm"),
    axis.ticks.length.y = unit(0.1, "cm"),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 12),
    strip.text = element_text(size = 15, hjust = 0, face = "italic")
  ) &
  NoLegend()
```

# Diff genes - His bundle
```{r}
VlnPlot(
  ohgoodyer_int,
  assay = "RNA",
  features = markers_diff_his,
  idents = c("P - His bundle", "E16.5 - His bundle"),
  fill.by = 'ident',
  stack = TRUE,
  flip = T,
  cols = c("#F6C141", "#F6C141")
) &
  stat_summary(
    fun = mean,
    geom = 'crossbar',
    width = 0.2,
    colour = "black"
  ) &
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    # axis.text.x = element_text(
    #     size = 16,
    #     angle = 40,
    #     vjust = 1,
    #     hjust = 1
    # ),
    axis.ticks.length.x = unit(0.1, "cm"),
    axis.ticks.length.y = unit(0.1, "cm"),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 12),
    strip.text = element_text(size = 15, hjust = 0, face = "italic")
  ) &
  NoLegend()
```

# Diff genes - PF
```{r}
VlnPlot(
  ohgoodyer_int,
  assay = "RNA",
  features = markers_diff_pf,
  idents = c("P - Purkinje fibers", "E16.5 - PF"),
  fill.by = 'ident',
  stack = TRUE,
  flip = T,
  cols = c("#512889", "#512889")
) &
  stat_summary(
    fun = mean,
    geom = 'crossbar',
    width = 0.2,
    colour = "black"
  ) &
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    # axis.text.x = element_text(
    #     size = 16,
    #     angle = 40,
    #     vjust = 1,
    #     hjust = 1
    # ),
    axis.ticks.length.x = unit(0.1, "cm"),
    axis.ticks.length.y = unit(0.1, "cm"),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 12),
    strip.text = element_text(size = 15, hjust = 0, face = "italic")
  ) &
  NoLegend()
```
