---
title: "R2R_sFig10_ionchannels"
output: R2R_sFig10_ionchannels
date: "2023-12-15"
editor_options: 
  chunk_output_type: console
---

# Obtain ion channel list - filtered (overlap between all genes in sc and spatial)
```{r}
ionchannel <- read.table("~/ionchannels_guidetopharmacology_org_nonzeroinsc.txt", sep = '\t', header = F)

ionchannel_update <- ionchannel$V1
ionchannel_update_voltage <- ionchannel_update[1:110]
ionchannel_update_ligand <- ionchannel_update[111:166]
ionchannel_update_others <- ionchannel_update[167:201]

```

# Voltage gated ionchannels - DotPlot
```{r}
DotPlot(
  ccs,
  features = ionchannel_update_voltage,
  cols = c('whitesmoke', 'deeppink4'),
  # col.min = 0,
  # col.max = 2,
  # dot.scale = 6,
  dot.min = -0.10
)  &
  guides(color = guide_colorbar(title = 'Average\nExpression')) &
  guides(size = guide_legend(title = 'Percent\nExpressed')) &
  RotatedAxis() &
  scale_size(breaks = c(0, 25, 50, 75, 100)) &
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
      size = 10,
      angle = 90,
      vjust = 0.5,
      hjust = 0.5
    ),
    axis.ticks.length.x.bottom = unit(3, 'mm'),
    axis.text.y = ggtext::element_markdown(size = 13,
                                           colour = rev(mycolors2)),
    legend.position = "right",
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.width = unit(0.35, 'cm'),
    legend.key.height = unit(0.35, 'cm')
  )
```

# Ligand gated ionchannels - DotPlot
```{r}
DotPlot(
  ccs,
  features = ionchannel_update_ligand,
  cols = c('whitesmoke', 'deeppink4'),
  dot.min = -0.10
)  &
  guides(color = guide_colorbar(title = 'Average\nExpression')) &
  guides(size = guide_legend(title = 'Percent\nExpressed')) &
  RotatedAxis() &
  scale_size(breaks = c(0, 25, 50, 75, 100)) &
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
      size = 10,
      angle = 90,
      vjust = 0.5,
      hjust = 0.5
    ),
    axis.ticks.length.x.bottom = unit(3, 'mm'),
    axis.text.y = ggtext::element_markdown(size = 13,
                                           colour = rev(mycolors2)),
    legend.position = "right",
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.width = unit(0.35, 'cm'),
    legend.key.height = unit(0.35, 'cm')
  )
```

# sFigx - Other ionchannels - DotPlot
```{r}
DotPlot(
  ccs,
  features = ionchannel_update_others,
  cols = c('whitesmoke', 'deeppink4'),
  dot.min = -0.10
)  &
  guides(color = guide_colorbar(title = 'Average\nExpression')) &
  guides(size = guide_legend(title = 'Percent\nExpressed')) &
  RotatedAxis() &
  scale_size(breaks = c(0, 25, 50, 75, 100)) &
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(
      size = 10,
      angle = 90,
      vjust = 0.5,
      hjust = 0.5
    ),
    axis.ticks.length.x.bottom = unit(3, 'mm'),
    axis.text.y = ggtext::element_markdown(size = 13,
                                           colour = rev(mycolors2)),
    legend.position = "right",
    legend.direction = "vertical",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.width = unit(0.35, 'cm'),
    legend.key.height = unit(0.35, 'cm')
  )
```



# Heatmap for spatial
```{r}
spatial_vgic <- read.table("~/spatial_guidetopharm_vgic_nozeroes_reorder.txt")
row.names(spatial_vgic) <- spatial_vgic$V1
spatial_vgic$V1 <- NULL
```

# Create labels
```{r}
spatialheatmap_labels <- read.table("q3_ExpressionFile_wlnb_forfig2_heatmap_groups.txt")
colnames(spatial_vgic) <- spatialheatmap_labels
```

# Scale data
```{r}
spatial_vgic_scaled <- as.matrix(t(scale(t(spatial_vgic))))
spatial_vgic_scaled <- t(spatial_vgic_scaled)
```

# Color palette - blue
```{r}
colorRampPalette(brewer.pal(9,"Blues"))(100)
heatmap_spatial_col = circlize::colorRamp2(c(-2, 0, 2), c("#F7FBFF", "#6DAFD6", "#08306B"))
```

# Fig2A - Spatial heatmap - Voltage gated ion channels
```{r}
tiff("~/spatial_vgic_reorder.tiff", width = 10, height = 2.6, units = "in", res = 500)

fig <- Heatmap(
  spatial_vgic_scaled,
  name = "Scaled\nExpression",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = heatmap_spatial_col,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 6),
  row_names_side = 'left',
  column_names_gp = gpar(fontsize = 6),
  column_names_side = 'top'
)

draw(fig)

dev.off()
```


# sFigx - Heatmap for spatial - Ligand gated ion channels
```{r}
spatial_lgic <- read.table("~/spatial_guidetopharm_ligand.txt")
row.names(spatial_lgic) <- spatial_lgic$V1
spatial_lgic$V1 <- NULL
```

# Create labels
```{r}
spatialheatmap_labels <- read.table("~/q3_ExpressionFile_wlnb_forfig2_heatmap_groups.txt")
colnames(spatial_lgic) <- spatialheatmap_labels
```

# Scale data
```{r}
spatial_ligand_scaled <- as.matrix(t(scale(t(spatial_lgic))))
spatial_ligand_scaled <- t(spatial_ligand_scaled)
```

# Color palette - blue
```{r}
colorRampPalette(brewer.pal(9,"Blues"))(100)
heatmap_spatial_col = circlize::colorRamp2(c(-2, 0, 2), c("#F7FBFF", "#6DAFD6", "#08306B")) 
```

# Fig2A - Spatial heatmap
```{r}
tiff("~/spatial_ligand.tiff", width = 8, height = 2.6, units = "in", res = 500)

fig <- Heatmap(
  spatial_ligand_scaled,
  name = "Scaled\nExpression",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = heatmap_spatial_col,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 6),
  row_names_side = 'left',
  column_names_gp = gpar(fontsize = 6),
  column_names_side = 'top'
)

draw(fig)

dev.off()
```




# Heatmap for spatial - Others
```{r}
spatial_others <- read.table("~/spatial_guidetopharm_others.txt")
row.names(spatial_others) <- spatial_others$V1
spatial_others$V1 <- NULL
```

# Create labels
```{r}
spatialheatmap_labels <- read.table("~/q3_ExpressionFile_wlnb_forfig2_heatmap_groups.txt")
colnames(spatial_others) <- spatialheatmap_labels
```

# Scale data
```{r}
spatial_others_scaled <- as.matrix(t(scale(t(spatial_others))))
spatial_others_scaled <- t(spatial_others_scaled)
```

# Color palette - blue
```{r}
colorRampPalette(brewer.pal(9,"Blues"))(100)
heatmap_spatial_col = circlize::colorRamp2(c(-2, 0, 2), c("#F7FBFF", "#6DAFD6", "#08306B"))
```

# Fig2A - Spatial heatmap
```{r}
tiff("~/spatial_others.tiff", width = 6, height = 3, units = "in", res = 500)

fig <- Heatmap(
  spatial_others_scaled,
  name = "Scaled\nExpression",
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  show_row_dend = FALSE,
  col = heatmap_spatial_col,
  use_raster = FALSE,
  row_names_gp = gpar(fontsize = 6),
  row_names_side = 'left',
  column_names_gp = gpar(fontsize = 6),
  column_names_side = 'top'
)

draw(fig)

dev.off()
```
